---
output:
  pdf_document: default
  html_document: default
---
Doina Covaliu - Gender Prediction using the first name
---

  
Install the necesarry packages (remove the comment sign before running the code)
```{r}
#install.packages("gender")
suppressPackageStartupMessages(library("gender"))
#install.packages("genderdata", repos = "https://dev.ropensci.org", type = "source")
suppressPackageStartupMessages(library("genderdata"))
#install.packages("dplyr")
suppressPackageStartupMessages(library("dplyr"))
#install.packages("stringr")
suppressPackageStartupMessages(library("stringr"))
#install.packages("tidyverse")
suppressPackageStartupMessages(library("tidyverse"))
#install.packages("magrittr")
suppressPackageStartupMessages(library("magrittr"))
#install.packages("DescTools")
suppressPackageStartupMessages(library("DescTools"))
```

Read the csv file into a dataframe
```{r}
data_source<-read.csv("C:/Users/Doina/Desktop/Customers_and_Orders.csv",
                      sep =",", stringsAsFactors = F)

```


```{r}
head(data_source)
tail(data_source)
str(data_source)
```
Data cleaning -Removing words like : INTERNATIONAL, REFUNDS, DELETE, comments that were included in some of the names.

```{r}
data_source<-dplyr::mutate_if(data_source, is.character,stringr::str_replace_all,
                              pattern="INTERNATIONAL ", replacement="")
data_source<-dplyr::mutate_if(data_source, is.character,stringr::str_replace_all,
                              pattern="REFUNDS ",replacement="")
data_source<-dplyr::mutate_if(data_source, is.character,stringr::str_replace_all,
                              pattern="DELETE ", replacement="")
data_source<-dplyr::mutate_if(data_source, is.character,stringr::str_replace_all,
                              pattern=" $", replacement="")
data_source<-dplyr::mutate_if(data_source, is.character,stringr::str_replace_all,
                              pattern="Estate of ", replacement="")
```

Remove "-" from the name
```{r}
data_source$First_Name<-str_replace_all(data_source$First_Name, "-","")

```

Remove space at the begining and the end of the name
```{r}
data_source$First_Name <- trimws(data_source$First_Name, which = c("both"))
```

Extract the data where the first name is a initial only,in a separate dataset.
In this case the gender cannot be determined .
```{r}
incomplete_names<-subset(data_source, grepl('^[a-zA-Z]{1}$|^[a-zA-Z]{1}\\.\\b|
^[a-zA-Z]{1}\\s{1}[A-Z]{1}\\b|^[a-zA-Z]{1}\\.\\s{1}[A-Z]{1}\\b
|^[a-zA-Z]{1}\\.\\s{1}[A-Z]{1}\\.\\b|^[a-zA-Z]{1}\\.[A-Z]{1}\\.\\b',
data_source$First_Name)==TRUE)
```


Removing the observations where the First Name is formed of initials.
```{r}
data_source<-subset(data_source, 
grepl('^[a-zA-Z]{1}$|^[a-zA-Z]{1}\\b|^[a-zA-Z]{1}$|^[a-zA-Z]{1}\\.\\b|
^[a-zA-Z]{1}\\s{1}[A-Z]{1}\\b|^[a-zA-Z]{1}\\.\\s{1}[A-Z]{1}\\b|
^[a-zA-Z]{1}\\.\\s{1}[A-Z]{1}\\.\\b|
^[a-zA-Z]{1}\\.[A-Z]{1}\\.\\b', data_source$First_Name)==FALSE)
```

Removing any "." from the names
```{r}
data_source$First_Name<-str_replace_all(data_source$First_Name, "\\.","")
```

For observation with the format: Name & Name, Name and Name the value "couple"
will be assigned to the Gender column
```{r}
data_source$Gender[data_source$First_Name %like any% c("% & %","% and %")]<-"couple"
couples<-subset(data_source, grepl("couple", data_source$Gender)==TRUE)
```

Temporary remove the couple from the dataset
```{r}
data_source<-subset(data_source, grepl("couple", data_source$Gender)==FALSE)
```

Remove the initial after the First Name or Before the first name
```{r}
data_source$First_Name<-str_replace_all(data_source$First_Name, "^[A-Z]{1}\\.\\s|^[A-Z]{1}\\s|\\s{1}[A-Z]{1}\\b$|\\s{1}[A-Z]{1}\\.\\b$|","")
```

If the First Names has 2 names remove the second name
```{r}
data_source$First_Name<-str_replace_all(data_source$First_Name, "\\s{1}[A-Za-z]{2,}$","")
```

Extract the unique first names from the dataset 
```{r}
names_unique<-unique(data_source[,1])
```

Predict gender and create a data frame of names & predicted genders
```{r}
predicted_names <- data.frame(gender(names_unique, method = "ssa"))
```

Assign the gender by joining the predicted_names dataset with the original dataset
```{r}
final_dataset<-left_join(data_source, predicted_names[,c(1,4)], by = c("First_Name" = "name"))
sapply(final_dataset, function(x) sum(is.na(x)))
#sapply(final_dataset2, function(x) sum(is.na(x)))


```
Remove the original gender column that has only n/a values and 
save the final dataset in a new csv file
```{r}
final_dataset<-final_dataset[-3]
final_dataset<-rename(final_dataset, Gender=gender)
final<-union(final_dataset, couples)
head(final)
tail(final)
#write.csv(final_dataset,"Final_dataset.csv")
#write.csv(final,"Final.csv")

```